---
output: github_document
---

```{r eval=FALSE, echo=FALSE}
# Build the paper:
Rscript -e 'rmarkdown::render("README.Rmd")'
```

# odjitter

This crate contains an implementation of the 'jittering' technique for pre-processing origin-destination (OD) data.
Jittering in a data visualisation context refers to the addition of "random noise to the data" to prevent points in graphs from overlapping, as described in by Wickham et al. ([2016](https://ggplot2-book.org/)) and in the documentation page for the function [`geom_jitter()`](https://ggplot2.tidyverse.org/reference/geom_jitter.html).

In the context of OD data jittering refers to randomly moving start and end points associated with OD pairs, as described in an under review paper on the subject (Lovelace et al. under review).
The technique is implemented in the function [`od_jitter()`](https://itsleeds.github.io/od/reference/od_jitter.html) in the [`od`](https://itsleeds.github.io/od/index.html) R package.
The functionality contained in this repo is an extended and much faster implementation: according to our benchmarks on a large dataset it was around 1000 times faster than the R implementation.

The crate is still a work in progress: the API may change.
Issues and pull requests are particularly useful at this stage.

<!-- Todo: add worked example. Will do for GISRUK paper -->
<!-- An illustration of its functionality is shown below. -->

# Installation

Install the package from the system command line as follows (you need to have installed and set-up [cargo](https://doc.rust-lang.org/cargo/getting-started/installation.html) first):

```bash
cargo install --git https://github.com/dabreegster/odjitter
```

To check the package installation worked, you can run `odjitter` command without arguments.
If it prints the following message congratulations, it works 🎉

```{r, engine='bash', error=TRUE}
odjitter
```

# Usage

To run algorithm you need a minimum of three inputs, examples of which are provided in the `data/` folder of this repo:

1. A .csv file containing OD data with two columns containing zone IDs (specified with  `--origin-key=geo_code1 --destination-key=geo_code2` by default) and other columns representing trip counts:
```{r, echo=FALSE, message=FALSE}
od = readr::read_csv("data/od.csv")
knitr::kable(od[1:3, ])
```
2. A [.geojson file](https://github.com/dabreegster/odjitter/blob/main/data/zones.geojson) representing zones that contains values matching the zone IDs in the OD data (the field containing zone IDs is specified with `--zone-name-key=InterZone` by default):
```{r, echo=FALSE}
# zones = sf::read_sf("data/zones.geojson")
# zones[1:3, ]
```
```{bash}
head -6 data/zones.geojson
```
3. A [.geojson file](https://github.com/dabreegster/odjitter/blob/main/data/road_network.geojson) representing a transport network from which origin and destination points are sampled
```{bash}
head -6 data/road_network.geojson
```

The `jitter` function requires you to set the maximum number of trips for all trips in the jittered result.
A value of 1 will create a line for every trip in the dataset, a value above the maximum number of trips in the 'all' column in the OD ata will result in a jittered dataset that has the same number of desire lines (the geographic representation of OD pairs) as in the input (50 in this case).

With reference to the test data in this repo, you can run the `jitter` command line tool as follows:

```{bash}
odjitter --od-csv-path data/od.csv \
  --zones-path data/zones.geojson \
  --subpoints-path data/road_network.geojson \
  --max-per-od 50 --output-path output_max50.geojson
```

Try running it with a different `max-per-od` value (10 in the command below):

```{bash}
odjitter --od-csv-path data/od.csv \
  --zones-path data/zones.geojson \
  --subpoints-path data/road_network.geojson \
  --max-per-od 10 --output-path output_max10.geojson
```

# Outputs

The figure below shows the output of the `jitter` commands above visually, with the left image showing unjittered results with origins and destinations going to zone centroids (as in many if not most visualisations of desire lines between zones), the central image showing the result after setting `max-per-od` argument to 50, and the right hand figure showing the result after setting `max-per-od` to 10.

```{r, echo=FALSE, fig.show='hold', out.width="30%", message=FALSE}
zones = sf::read_sf("data/zones.geojson")
network = sf::read_sf("data/road_network.geojson")
od_sf = od::od_to_sf(od, zones)
plot(zones$geometry, border = "grey")
plot(network$geometry, col = "red", add = TRUE)
plot(od_sf$geometry, add = TRUE)
plot(zones$geometry, border = "grey")
plot(network$geometry, col = "red", add = TRUE)
plot(sf::read_sf("output_max50.geojson")$geometry, add = TRUE)
plot(zones$geometry, border = "grey")
plot(network$geometry, col = "red", add = TRUE)
plot(sf::read_sf("output_max10.geojson")$geometry, add = TRUE)
```

Note: `odjitter` uses a random number generator to sample points, so the output will change each time you run it, unless you set the `rng-seed`, as documented in the next section.

# Details

For full details on `odjitter`'s arguments run `odjitter --help` which gives the following output:

```{bash}
odjitter --help
```

# References

Lovelace, Robin, Rosa Félix, and Dustin Carlino Under Review Jittering: A Computationally Efficient Method for Generating Realistic Route Networks from Origin-Destination Data. TBC.

Wickham, Hadley 2016 ggplot2: Elegant Graphics for Data Analysis. 2nd ed. 2016 edition. New York, NY: Springer.