---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# odr

<!-- badges: start -->
<!-- badges: end -->

The goal of {odrust} is to provide an R interface to the [odjitter](https://github.com/dabreegster/odjitter) Rust crate for processing origin-destination data.

```{r, eval=FALSE}
remotes::install_github("atumworld/odrust")
```

```{r}
library(odrust)
```


## R interface to `odjitter` Rust crate via system commands


```{r jitter, out.width="50%", fig.show='hold'}
od = readr::read_csv("https://github.com/dabreegster/odjitter/raw/main/data/od.csv")
zones = sf::read_sf("https://github.com/dabreegster/odjitter/raw/main/data/zones.geojson")
names(zones)[1] = "geo_code"
road_network = sf::read_sf("https://github.com/dabreegster/odjitter/raw/main/data/road_network.geojson")
od_unjittered = od::od_to_sf(od, zones)
set.seed(42) # for reproducibility
od_jittered = odr_jitter(od, zones, subpoints = road_network)
nrow(od_unjittered)
nrow(od_jittered)
plot(od_unjittered)
plot(od_jittered)
```

## R interface to Rust via rextendr (not currently working)

The development of the package was done using the development version of the `rextendr` package.

```{r, eval=FALSE}
remotes::install_github("extendr/rextendr")
```

The package template was created as follows:

```{r, eval=FALSE}
usethis::use_description()
rextendr::use_extendr()
```

The odjitter Rust crate ported into the src/rust folder.

```{r, eval=FALSE, echo=FALSE}
list.files("src/rust/")
file.edit("~/orgs/atumWorld/odjitter/src/lib.rs")
file.edit("~/orgs/atumWorld/odjitter/Cargo.toml")
file.edit("src/rust/Cargo.toml")
```


```{bash, echo=FALSE, eval=FALSE}
cp -Rv ~/orgs/atumWorld/odjitter/src/scrape.rs src/rust/src/
```

